Manuscript draft

### 2019-06-07 manuscript/manuscript draft.Rmd
updated manuscript draft

---
title: "Manuscript draft"
output:
  pdf_document:
    citation_package: natbib
bibliography: library.bib
csl: apa.csl
---

, fig.cap = "Map showing all data used in the analysis, coded by variable"

@@

Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018]. This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from previously published studies and existing databases (e.g. ORNL DAAC; Luyssaert etc?). Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.

  ```{r echo=FALSE, out.width='100%'}
  knitr::include_graphics("distribution_all_variables.png")
  ```

*Data selection.* ForC contains measures of carbon stocks and fluxes; for the purposes of this analysis only measures of primary productivity were selected from the database. The variables selected were: GPP, NPP, BNPP (total root and fine root), ANPP (total ANPP, foliage, wood and woody stem).

| Variable | Definition | Components included | Methodologies used|
|-------------|-------------------|-------------|------------------------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and optionally understory | Direct measurement of annual increments of components | 
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 
| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |

Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. Sites that had experienced significant disturbance were also excluded. Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.  
There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015]. To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.

*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods. For this reason, criteria were introduced to standardise for differences in methodology. Where data was based on forest plot census measurements, studies which uses a minimum dbh measure of 10cm or greater were excluded from analysis.  
Estimates of total ecosystem primary productivity are based on measurements of the component parts of forest productivity. Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis. Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and understory. Measures of NPP which included additional components were excluded. Estimates of ANPP were selected if they included foliage, stem growth and branch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.

*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database. In addition to this data, climate data for each site was extracted from five open-access climate datasets.

@@

>>>>

---
title: "Manuscript draft"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
bibliography: library.bib
csl: apa.csl
header-includes:  \usepackage{float}
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

@@

Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018]. This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ??? Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions. ForC currently contains [] records from [] plots, representing [] distinct geographic areas across all forested biogeographic and climate zones.

  ```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
  knitr::include_graphics("distribution_all_variables.png")
  ```
*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity. Table 1 contains details of the variables selected for analysis.

| Variable | Definition | Components included | Methodologies used|
|--------------|-------------------|-------------|------------------------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and optionally understory | Direct measurement of annual increments of components | 
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 
| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |

Table: Definitions of variables used in analysis

A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors. Since management can alter observed patterns of primary productivity [@Simova2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. Sites that had experienced significant disturbance were also excluded. Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.  
There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015]. To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis. Sites for which stand age was unknown were excluded from analysis.

*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods. For this reason, criteria were introduced to standardise for differences in methodology. Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of 10cm or greater were excluded from analysis.  
Estimates of total ecosystem primary productivity are based on measurements of the component parts of forest productivity. Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis. Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and understory. Measures of NPP which included additional components were excluded. Estimates of ANPP were selected if they included foliage, stem growth and branch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.

*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database. In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.

@@

Table: Sources of climate data
